Hierarchical Reverberation Mapping

### 2013-10-29 Paper/ms.tex
More "scaffolding" type writing

\section{Combining Inferences About Multiple Objects}\label{sec:multiple}
Consider a sample of $N$ objects, each of which has parameters $\theta_i$.
If the $i$th object is analysed on its own, the inference about 
its parameters $\theta_i$ is described by a posterior distribution
\begin{eqnarray}
p(\theta_i | x_i) \propto \pi(\theta_i)p(x_i | \theta_i)\label{eq:individual}
\end{eqnarray}
where $\pi(\theta_i)$ is the prior distribution and $p(x_i | \theta_i)$ is the
likelihood function. If $N$ objects are analysed separately, inferences can
be made about the diversity of $\theta_i$ across the sample. However, these
inferences may be incorrect due to the implicit assumption that the prior for
all of the $\{\theta_i\}$ is independent.
A hierarchical model may be used to overcome this problem.
Section~\ref{sec:single_object} describes the assumptions used for analysing
data from a single AGN.

In a hierarchical model, the prior for the parameters $\{\theta_i\}$ is created
by introducing hyperparameters $\alpha$ such that the joint prior for $\alpha$
and the $\{\theta_i\}$ is:
\begin{eqnarray}
p(\alpha, \theta_i) &=& p(\alpha)\prod_{i=1}^N p(\theta_i | \alpha)
\end{eqnarray}
The posterior distribution for
$\btheta = \{\theta_1, ..., \theta_N\}$ and the hyperparameters $\alpha$
given the data 
$\bx = \{x_1, ..., x_N\}$
should actually be
\begin{eqnarray}
p(\alpha, \btheta | \bx) &\propto&
p(\alpha)p(\btheta|\alpha)p(\bx | \btheta, \alpha)\\
&=& p(\alpha)\prod_{i=1}^N f(\theta_i|\alpha)p(x_i | \theta_i)
\end{eqnarray}
The marginal posterior distribution for the hyperparameters $\alpha$ is
\begin{eqnarray}
p(\alpha | \bx) &=&
\int p(\alpha, \btheta|\bx) \, d\btheta\\
&\propto& p(\alpha)\int \prod_{i=1}^N f(\theta_i|\alpha)p(x_i | \theta_i) d^N\theta\\
&\propto& p(\alpha) \prod_{i=1}^N \int f(\theta_i|\alpha)p(x_i | \theta_i) d\theta_i\\
&\propto& p(\alpha) \prod_{i=1}^N \int \frac{f(\theta_i|\alpha)}{\pi(\theta)}p(x_i | \theta_i) \pi(\theta_i)d\theta_i\\
&\propto& p(\alpha) \prod_{i=1}^N \mathds{E}\left[\frac{f(\theta_i|\alpha)}{\pi(\theta)}\right]
\end{eqnarray}
where the expectation is taken with respect to the individual object posterior
of Equation~\ref{eq:individual}, and thus can be estimated using posterior
samples from the posterior distributions for the individual objects.
This result enables us to reconstruct the
posterior distribution for the hyperparameters even though the individual object
inferences were made without the hierarchical structure in the prior. This is
essentially an importance sampling approximation to the hierarchical model.

Rather than
analysing all objects together, which can be very computationally expensive,
we can use the posterior distributions from Equation~\ref{eq:individual}
to reconstruct the results that would be obtained if we did analyse all
objects together with a hierarchical prior on the $\theta_i$ parameters.

\section{The Single Object Model}\label{sec:single_object}

\section{Model Assumptions}
If the continuum light curve of an AGN
is described by a function $y(t)$, then the line
light curve $l(t)$ is given by
\begin{eqnarray}
l(t) &=& A \int_\tau \Psi(\tau)\left[y(t - \tau) + C\right] \, d\tau
\end{eqnarray}
where $\Psi(\tau)$ is the transfer function (assumed to be normalised),
and $A$ and $C$ are response
coefficients. The idea of reverberation mapping is to use noisy measurements
of $y(t)$ and $l(t)$ to infer the transfer function $\Psi(\tau)$ or a summary
of it such as the mean lag $\bar{\tau} = \int \tau\Psi(\tau) \, d\tau$.
Throughout this letter we consider $y(t)$ and $l(t)$ in flux
units, as opposed to magnitudes.

The posterior distribution for the mean lag $\bar{\tau}$ of a single object
$i$ may be obtained
by fitting the following model. We assume, for simplicity, that
the transfer function is uniform
between limits $a$ and $b$, where $b > a$:
\begin{eqnarray}
\Psi(\tau) &=& \left\{
\begin{array}{lr}
\frac{1}{b-a}, & \tau \in [a,b]\\
0, & \textnormal{otherwise}s
\end{array}\right.
\end{eqnarray}
and our goal is to measure the mean lag:
\begin{eqnarray}
\bar{\tau} &=& \int \tau \Psi(\tau) \, d\tau\\
&=& \frac{1}{2}(b-a)
\end{eqnarray}

>>>>

\section{Combining Inferences About Multiple Objects}\label{sec:multiple}
Consider a sample of $N$ objects, each of which has parameters $\theta_i$.
If the $i$th object is analysed on its own, the inference about 
its parameters $\theta_i$ is described by a posterior distribution
\begin{eqnarray}
p(\theta_i | x_i) \propto \pi(\theta_i)p(x_i | \theta_i)\label{eq:individual}
\end{eqnarray}
where $\pi(\theta_i)$ is the prior distribution and $p(x_i | \theta_i)$ is the
likelihood function. If $N$ objects are analysed separately, inferences can
be made about the diversity of $\theta_i$ across the sample. However, these
inferences may be incorrect due to the implicit assumption that the prior for
all of the $\{\theta_i\}$ is independent \citep{2013arXiv1310.5177B}.
A hierarchical model may be used to overcome this problem.
Section~\ref{sec:single_object} describes the assumptions used for analysing
data from a single AGN.

In a hierarchical model, the prior for the parameters $\{\theta_i\}$ is created
by introducing hyperparameters $\alpha$ such that the joint prior for $\alpha$
and the $\{\theta_i\}$ is:
\begin{eqnarray}
p(\alpha, \theta_i) &=& p(\alpha)\prod_{i=1}^N p(\theta_i | \alpha)
\end{eqnarray}
The implied prior on the $\{\theta_i\}$ parameters is then
\begin{eqnarray}
p(\theta_i) &=& \int p(\alpha)\prod_{i=1}^N p(\theta_i | \alpha) \, d\alpha
\end{eqnarray}
which may imply that the $\theta$ values are ``clustered'' around some
typical value.

The posterior distribution for
$\btheta = \{\theta_1, ..., \theta_N\}$ and the hyperparameters $\alpha$
given the data 
$\bx = \{x_1, ..., x_N\}$, is then
\begin{eqnarray}
p(\alpha, \btheta | \bx) &\propto&
p(\alpha)p(\btheta|\alpha)p(\bx | \btheta, \alpha)\\
&=& p(\alpha)\prod_{i=1}^N f(\theta_i|\alpha)p(x_i | \theta_i)
\end{eqnarray}

It is possible to consider the data for all $N$ objects as one big data set
and infer the parameters of all objects simultaneously. However, this can be
computationally prohibitive. In some circumstances it is tractable to analyse
each object's data separately (using a common prior $\pi(\theta_i)$) and
then post-process the results, reconstructing what the results of the
hierarchical model would have been. In this letter we use this approach.

The marginal posterior distribution for the hyperparameters $\alpha$ is
\begin{eqnarray}
p(\alpha | \bx) &=&
\int p(\alpha, \btheta|\bx) \, d\btheta\\
&\propto& p(\alpha)\int \prod_{i=1}^N f(\theta_i|\alpha)p(x_i | \theta_i) d^N\theta\\
&\propto& p(\alpha) \prod_{i=1}^N \int f(\theta_i|\alpha)p(x_i | \theta_i) d\theta_i\\
&\propto& p(\alpha) \prod_{i=1}^N \int \frac{f(\theta_i|\alpha)}{\pi(\theta)}p(x_i | \theta_i) \pi(\theta_i)d\theta_i\\
&\propto& p(\alpha) \prod_{i=1}^N \mathds{E}\left[\frac{f(\theta_i|\alpha)}{\pi(\theta)}\right]\label{eq:combine}
\end{eqnarray}
where the expectation is taken with respect to the individual object posterior
of Equation~\ref{eq:individual}, and thus can be estimated using posterior
samples from the posterior distributions for the individual objects.
This result enables us to reconstruct the
posterior distribution for the hyperparameters even though the individual object
inferences were made without the hierarchical structure in the prior. This is
essentially an importance sampling approximation to the
full hierarchical model.

In Section~\ref{sec:single_object} we define a simple model for inferring the
lag of a single AGN from sparse RM data. We will use Markov Chain Monte Carlo
(MCMC) to produce posterior samples for the lags of the $N$ objects, which can
then be combined using Equation~\ref{eq:combine} to yield the posterior
distribution for some hyperparameters describing the distribution of
lags in the sample.

\section{The Single Object Model}\label{sec:single_object}
If the continuum light curve of an AGN
is described by a function $y(t)$, then the line
light curve $l(t)$ is given by
\begin{eqnarray}
l(t) &=& A \int_\tau \Psi(\tau)\left[y(t - \tau) + C\right] \, d\tau
\end{eqnarray}
where $\Psi(\tau)$ is the transfer function (assumed to be normalised),
and $A$ and $C$ are response
coefficients. The idea of reverberation mapping is to use noisy measurements
of $y(t)$ and $l(t)$ to infer the transfer function $\Psi(\tau)$ or a summary
of it such as the mean lag $\bar{\tau} = \int \tau\Psi(\tau) \, d\tau$.
Throughout this letter we consider $y(t)$ and $l(t)$ in flux
units, as opposed to magnitudes, and consider $\bar{\tau}$ as
the definition of ``the lag'' of the AGN.

The posterior distribution for the lag $\bar{\tau}$ of a single object
$i$ may be obtained
by fitting the following model. We assume, for simplicity, that
the transfer function is uniform
between limits $a$ and $b$, where $b > a$:
\begin{eqnarray}
\Psi(\tau) &=& \left\{
\begin{array}{lr}
\frac{1}{b-a}, & \tau \in [a,b]\\
0, & \textnormal{otherwise}s
\end{array}\right.
\end{eqnarray}
and our goal is to measure the mean lag:
\begin{eqnarray}
\bar{\tau} &=& \int \tau \Psi(\tau) \, d\tau\\
&=& \frac{1}{2}(b-a)
\end{eqnarray}
